---
title: "Classic examples"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r load, warning=FALSE, message=FALSE}
library(diffuseR)
library(scatterplot3d)
library(diffusionMap)
```
A few other examples that demonstrate the performance of diffusion maps with randomized SVD against traditional diffusion maps. 

# Trefoil knot 

A nontrival knot that we generate as shown below. 

```{r, message=FALSE, warning=FALSE}
t <- runif(1000, 0, 2*pi)
x <- (2 + cos(3*t))*cos(2*t)
y <- (2 + cos(3*t))*sin(2*t)
z <- sin(3*t)

# Viz 
scatterplot3d(x, y, z, 
              highlight.3d = T,
              xlab = NULL, 
              ylab = NULL, 
              main = NULL, 
              grid = FALSE,
              tick.marks = FALSE,
              axis = FALSE)
```

As shown the diffusion coordinates from our adapted diffusion map algorithm with randomized SVD does not unravel this knot as well as that of the traditional mapping: 

```{r, message=FALSE, warning=FALSE}
D <- as.matrix(dist(cbind(x, y, z)))
trefoil_new <- diffuseR::diffusionmapRsvd(D)$X
trefoil_old <- diffusionMap::diffuse(D)$X
# For viz
rbPal <- colorRampPalette(c('red','black'))
col_new <- rbPal(15)[as.numeric(cut(trefoil_new[,1],breaks = 15))]
col_old <- rbPal(15)[as.numeric(cut(trefoil_old[,1],breaks = 15))]
# Plot 
par(mfrow = c(1,2))
plot(trefoil_new[,1], trefoil_new[,2], pch = 20, col = col_new, axes = FALSE, ann = FALSE)
plot(trefoil_old[,1], trefoil_old[,2], pch = 20, col = col_old, axes = FALSE, ann = FALSE)
```

# Chainlink

Here's another classical dataset from the $\texttt{diffusionMap}$ package where there are clearly two different classes, but they are entangled. 

```{r, message=FALSE, warning=FALSE}
data("Chainlink")
lab.col <- c(rep("red",500), rep("black",500))
scatterplot3d(Chainlink$C1, Chainlink$C2, Chainlink$C3,
              color = lab.col,
              pch = 20,
              xlab = NULL, 
              ylab = NULL, 
              main = NULL, 
              grid = FALSE,
              tick.marks = FALSE,
              axis = FALSE) 
```

When we apply our adapted diffusion map, again see the same issues: 

```{r, message=FALSE, warning=FALSE}
D <- as.matrix(dist(Chainlink))
chain_new <- diffuseR::diffusionmapRsvd(D)$X
chain_old <- diffusionMap::diffuse(D)$X
# For viz
rbPal <- colorRampPalette(c('red','black'))
col_new <- rbPal(15)[as.numeric(cut(chain_new[,1],breaks = 15))]
col_old <- rbPal(15)[as.numeric(cut(chain_old[,1],breaks = 15))]
# Plot 
par(mfrow = c(1,2))
plot(chain_new[,1], chain_new[,2], pch = 20, col = col_new, axes = FALSE, ann = FALSE)
plot(chain_old[,1], chain_old[,2], pch = 20, col = col_old, axes = FALSE, ann = FALSE)
```

# Swiss roll

This classic dataset is relied upon for testing various dimensionality reduction algos. 

```{r, message=FALSE, warning=FALSE}
n <- 800
t <- (3*pi/2)*(1+2*runif(n))
X <- cbind(t*cos(t), runif(n), t*sin(t))
scatterplot3d(X, highlight.3d = TRUE, 
              pch = 20,
              xlab = NULL, 
              ylab = NULL, 
              main = NULL, 
              grid = FALSE,
              tick.marks = FALSE,
              axis = FALSE)
```

Again, the differences between utilizing our adapted diffusion map algorithm against that provided by the $\texttt{diffusionMap}$ package: 

```{r, message=FALSE, warning=FALSE}
D <- as.matrix(dist(X))
swiss_new <- diffuseR::diffusionmapRsvd(D)$X
swiss_old <- diffusionMap::diffuse(D)$X
# For viz
rbPal <- colorRampPalette(c('red','black'))
col_new <- rbPal(15)[as.numeric(cut(swiss_new[,1],breaks = 15))]
col_old <- rbPal(15)[as.numeric(cut(swiss_old[,1],breaks = 15))]
# Plot 
par(mfrow = c(1,2))
plot(swiss_new[,1], swiss_new[,2], pch = 20, col = col_new, axes = FALSE, ann = FALSE)
plot(swiss_old[,1], swiss_old[,2], pch = 20, col = col_old, axes = FALSE, ann = FALSE)
```

# Annulus 

Another classical example utilized in convex analysis is the annulus data set. It consists of a region bounded by two circles like so:

```{r, message=FALSE, warning=FALSE}
data("annulus")
X <- annulus
plot(X, axes = F, pch = 20, ann = F)
```

```{r, message=FALSE, warning=FALSE}
D <- as.matrix(dist(X))
ann_new <- diffuseR::diffusionmapRsvd(D)$X
ann_old <- diffusionMap::diffuse(D)$X
# For viz
rbPal <- colorRampPalette(c('red','black'))
col_new <- rbPal(15)[as.numeric(cut(ann_new[,1],breaks = 15))]
col_old <- rbPal(15)[as.numeric(cut(ann_old[,1],breaks = 15))]
# Plot 
par(mfrow = c(1,2))
plot(ann_new[,1], ann_new[,2], pch = 20, col = col_new, axes = FALSE, ann = FALSE)
plot(ann_old[,2], ann_old[,3], pch = 20, col = col_old, axes = FALSE, ann = FALSE)
```